\toolnameend: A Tool for Visualizing Patterns of Spreadsheet Function Combinations

### 2016-08-11 doc/paper.tex
Introduction revisions

	\section{Introduction} Spreadsheets surround us. They serve key roles in
	industry, where as much as 95\% of U.S. financial firms use them
	daily~\cite{panko2008sarbanes}, and academia, where teachers can use them to
	track research and students' grades, not to mention many other
	applications~\cite{ko2011state}. It's this flexibility that grants them their
	allure: while a novice can still work without much experience, spreadsheets
	offer hundreds of specialized functions to fulfill a wide range of advanced
	needs, from complex statistics to text
	manipulation~\cite{nardi1990spreadsheet}. As such, it should be no surprise
	that tens of millions of workers rely on these programs
	daily~\cite{scaffidi2005estimating}.

	However, this ubiquity is not without danger; as more people adopt these tools
	and make larger and more complex sheets, the cost of failure grows too. To
	illustrate this, the European Spreadsheet Risks Interest Group (EuSpRIG)
	documents many horror stories of spreadsheets gone
	wrong\footnote{http://www.eusprig.org/horror-stories.htm}. One story tells of a
	2013 economics paper which drew connections between debt and national growth.
	Its findings shaped many political initiatives throughout the United States and
	Europe. However, other researchers soon discovered that a critical selection
	error in the research spreadsheets hampered the results and that, by correcting
	this error, the findings reversed completely! And this is only one of several
	accounts; even if most spreadsheet errors are relatively benign, poor
	spreadsheet practice can risk millions, if not billions, of
	dollars~\cite{powell2009impact}.

	From these stories, we can see the dire importance of understanding how people
	actually use spreadsheets. Fortunately, spreadsheet researchers have supported
	this by amassing spreadsheet collections from diverse sources, industrial to
	academic. Collections like EUSES~\cite{fisher2005euses},
	Fuse~\cite{barik2015fuse}, and the Enron corpus~\cite{hermans2015enron} have
	already fueled productive research across the field, such as in the search for
	code smells in spreadsheet
	environments~\cite{hermans2012detecting}~\cite{jansen2015code} or the creation
	of grammars to capture every possible spreadsheet program in
	Excel~\cite{aivaloglou2015grammar}. Additionally, some research has focused on
	comparing these collections, as Jansen did in 2015, to discover variation in
	spreadsheet techniques over time and location~\cite{jansen2015enron}.

	Considering the value of spreadsheets and all of these resources to describe
	their use, it becomes necessary that we design tools to better enable the
	exploration of these datasets and simplify the complex concepts that occur in
	spreadsheet design. One such concept is that of formula construction; functions
	are an important part of using spreadsheets efficiently, yet the number of
	potential combinations between them quickly becomes massive when you consider
	that there are hundreds of them. Tools, then, that empower people to explore
	questions about which functions are used together in practice have the
	potential to improve the current design of spreadsheets themselves and educate
	new generations of spreadsheets users on how to work best.

	This paper contributes such a tool, \toolnameend, that attempts to fulfill this
	need for better exploration by building on the aforementioned spreadsheet
	corpora. Working with spreadsheets in Excel, \toolname focuses primarily on the
	functions, like SUM and IF and hundreds more, which are the building blocks of
	formulae that determine values across a spreadsheet. The tool creates an
	interactive, exploratory visualization which captures how people combine these
	functions in a selected collection of spreadsheets, quantifying the broad
	patterns of function nesting while also being specific enough to find anomalous
	formulae and link them back to the specific cell in spreadsheets where they
	originate. \par

	Alongside the tool, we discuss several of the key decisions we faced during
	design, such as what to do with redundant formulae in a sheet or how to
	represent optional arguments, and we describe how our overall design goals both
	helped and hindered this process. We then apply the tool to a number of case
	studies to evaluate how much it could contribute to several different contexts,
	most notably code smell detection and spreadsheet education. To better
	understand how people might use \toolnameend, we also conducted a user study to
	judge how usable and intuitive the tool actually is. This grounds the
	discussion of its potential uses while also underscoring some of its inherent
	limitations and setting the way for future extensions and improvements upon the
	concept.

>>>>

	\section{Introduction} Spreadsheets surround us. They are vital to
	industry and academia alike, supporting as much as 95\% of U.S. financial firms use them
	daily~\cite{panko2008sarbanes} and empowering teachers to
	track research and students' grades~\cite{ko2011state}. This flexibility boosts their
	allure: while a novice can work with basic offerings, advanced users can choose
	among the hundreds of specialized functions to fulfill a wide range of
	needs, from complex statistics to text
	manipulation~\cite{nardi1990spreadsheet}. As such, it should be no surprise
	that tens of millions of workers rely on these programs
	daily~\cite{scaffidi2005estimating}.

	However, this ubiquity is not without danger; as more people make larger and
	more complex spreadsheets, the cost of failure grows too. The European
	Spreadsheet Risks Interest Group (EuSpRIG) documents many horror stories of
	spreadsheets gone wrong\footnote{http://www.eusprig.org/horror-stories.htm},
	where consequences range from the profuse loss of money to the inaccurate
	outcomes of elections to the creation of extensive political initiatives based
	on flawed economics. Even if most spreadsheet errors are relatively benign,
	poor spreadsheet practice can risk millions, if not billions, of
	dollars~\cite{powell2009impact}.

	Fortunately, spreadsheet researchers address these dangers by amassing
	spreadsheet collections from diverse sources, industrial to academic, so
	researchers can study how people use spreadsheets. Collections like
	EUSES~\cite{fisher2005euses}, Fuse~\cite{barik2015fuse}, and the Enron
	corpus~\cite{hermans2015enron} have already produced valueable research, such as the
	documentation of code smells in spreadsheet
	environments~\cite{hermans2012detecting}~\cite{jansen2015code} or the creation
	of grammars to describe every possible spreadsheet program in
	Excel~\cite{aivaloglou2015grammar}. Additionally, some research compares these
	collections to discover variation in spreadsheet techniques over time and
	location~\cite{jansen2015enron}.

	The value of spreadsheets demands that we design tools to better understand
	these datasets and simplify complexities we find in them. One such complexity
	is formula construction; functions are an important part of using spreadsheets
	efficiently, yet the number of potential combinations is massive considering
	that Excel boasts over 400 functions. Tools that explore these combinations in
	practice can aid researchers in understanding how people program and in
	educating new generations of spreadsheets users on how to work best.

	This paper contributes such a tool, \toolnameend, that attempts to fulfill this
	need for better exploration of spreadsheet
	corpora. Working with spreadsheets in Excel, \toolname focuses primarily on the
	functions, like SUM and IF and hundreds more, which are the building blocks of
	formulae that determine values across a spreadsheet. The tool creates an
	interactive visualization which captures how people combine these
	functions, quantifying the broad
	patterns of function nesting while also capturing anomalous
	formulae and their locations.

	We also discuss several of the key decisions we faced during
	design, such as what to do with a sheet's redundant formulae or how to
	represent optional arguments, and we describe how our overall design goals both
	helped and hindered this process. We then apply the tool to a number of case
	studies to evaluate how much it could contribute to several different contexts,
	most notably code smell detection and spreadsheet education. To better
	understand how people might use \toolnameend, we also conducted a user study to
	judge how usable and intuitive the tool actually is. This grounds the
	discussion of its potential uses while also underscoring some of its inherent
	limitations and setting the way for future extensions and improvements upon the
	concept.
